Modelling the emergence of whisker barrels

### 2019-12-17 paper/briefpaper/brief.tex
Updated figure (corrected black boundaries between regions and unified font sizes)

\caption{\textbf{C.} Cortical whisker barrels from \cite{zheng_signal_2001}
  with annotations for approximate axes and lines drawn to join the
  interstices into a putative Voronoi diagram.  \textbf{B.}  Top: Stochastic
  initial condition for $a_i$ with $i=13$ (barrel C4). The roll-off towards
  the boundary (thin line) ensures the boundary conditions are
  satisfied. Bottom: Guidance molecules in the simulated cortex. Molecule 1 is
  expressed high to low in an approximately anterior to posterior
  gradient. Molecule 2 is expressed lateral to medial (analogously to ephrin -
  Vanderhaegen et al) \textbf{C.}  Barreloid centres for mystacial whiskers
  taken from \cite{haidarliu_size_2001} are shown as black circles. The dashed
  grey line shows the boundary of the barreloid field. It is assumed that
  there are two linearly increasing, orthogonal signalling molecules in the
  thalamus, such that the location of each barreloid within the graded
  thalamic signal field can differentiate the efferent axons, with the result
  that axons from each barreloid interact to varying degrees with the cortical
  signalling gradients shown in B. Coloured arrows show, for each barreloid,
  the magnitude (and sign) of the interaction that its thalamocortical
  projection will have with each signalling gradient in the cortex. The
  mappings from barreloid location to interaction strength are
  linear. \textbf{D.}  Barrel cortex simulation. Colours indicate
  $\arg\max_i(c_i)$ for 41 is. Parameters:
  $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$,
  $\delta{t}=0.0001$. \textbf{E.}  The Honda Dirichlet measure, $\delta$, as a
  function of simulation time for the simulations show in D. \textbf{F.}
  Guidance molecules in the simulated Fgf8 misexpression experiment. Molecule
  1 is mirrored over the anterior-posterior axis The simulation indicates that
  the mirroring of the barrel field can occur as a result of the mirroring of
  guidance molecule 1.}
\label{fig:main}
\end{figure*}

First we verified that all results established by \cite{karbowski_model_2004} for a 1D axis could be reproduced using our extension to a 2D cortical sheet. Using an elliptical boundary with $M=3$ offset guidance gradients aligned to the longer axis, $N=5$ thalamocortical projections gave rise to five distinct cortical fields at locations that preserved the topographic ordering defined by the original $\gamma$ values. However, we found that specifying $N$ ordered areas required $M\approx N$ signalling fields. As the number of distinct guidance fields is unlikely to approach the number of barrels, further extensions were required.

>>>>

\caption{\textbf{A.} Cortical whisker barrels from \cite{zheng_signal_2001}
  annotated with lines joining interstices into a putative Voronoi diagram
  ($\delta=0.022$). \textbf{B.}  Top: Stochastic initial condition for $a_i$
  with $i=1$ (barrel `$\alpha$'). The roll-off towards the boundary (thin line)
  ensures the boundary conditions are satisfied. Bottom: Guidance molecules
  $\rho_1(\mathbf{x})$ and $\rho_2(\mathbf{x})$ in the simulated
  cortex. $\rho_1(\mathbf{x})$ varies along the anterior-posterior axis;
  $\rho_2(\mathbf{x})$ is expressed along the lateral/medial axis, analogous
  to ephrin A5 \citep{vanderhaeghen_mapping_2000}. \textbf{C.}  Barreloid
  centres for mystacial whiskers taken from \cite{haidarliu_size_2001} are
  shown as black circles. The dashed grey line shows the boundary of the
  barreloid field. It is assumed that there are two linearly increasing,
  orthogonal signalling molecules in the thalamus, such that the location of
  each barreloid within the graded thalamic signal field can differentiate the
  efferent axons, with the result that axons from each barreloid interact to
  varying degrees with the cortical signalling gradients $\rho_1$ and
  $\rho_2$. Blue (green) arrows show, for each barreloid, the magnitude (and
  sign) of the interaction that its thalamocortical projection will have with
  $\rho_1$ ($\rho_2$). The mappings from barreloid location to interaction
  strength are linear. \textbf{D.}  Barrel cortex simulation with parameters
  $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$ and
  $\delta{t}=0.0001$. Colours indicate $\arg\max_i(c_i)$ for 41 TC
  projections.  \textbf{E.}  The Honda Voronoi measure, $\delta$, (red) and a
  pattern fitting metric based on the distance, $d_i$, between simulated and
  experimental barrel centroids and the absolute difference in their areas,
  $\Delta A_i$ given by $\sigma = \sum_i^N d_i^2 \times \sum_i^N \Delta A_i$
  (black) plotted as a function of time for the simulations shown in
  D. \textbf{F.}  Guidance molecules in the simulated Fgf8 misexpression
  experiment. $\rho_1$ is mirrored over the anterior-posterior axis. The
  simulation indicates that the mirroring of the barrel field can occur as a
  result of the mirroring of guidance molecule 1.}
\label{fig:main}
\end{figure*}

First we verified that all results established by \cite{karbowski_model_2004} for a 1D axis could be reproduced using our extension to a 2D cortical sheet. Using an elliptical boundary with $M=3$ offset guidance gradients aligned to the longer axis, $N=5$ thalamocortical projections gave rise to five distinct cortical fields at locations that preserved the topographic ordering defined by the original $\gamma$ values. However, we found that specifying $N$ ordered areas required $M\approx (N+1)/2$ signalling fields. As the number of distinct guidance fields is unlikely to approach the number of barrels, further extensions were required.
